Deciding Bitvector Termination with Program Synthesis

### 2014-06-23 papers/termination/paper.tex
Discussion of linearity vs. LKC

\begin{definition}[Synthesis Fragment]
 A formula is in the synthesis fragment iff it is of the form
 \[
  \exists P, x \cdot \forall y \cdot \sigma(P, x, y)
 \]

@@

\begin{figure*}
\begin{framed}
\begin{definition}[Second-order Termination Formula]
\label{def:termination-formula}
\begin{align*}
 \exists R \cdot \forall x, x' \cdot & L(x) \wedge T(x, x') \rightarrow R(x) > 0 \wedge R(x) > R(x')
\end{align*}
\end{definition}

\begin{definition}[Conditional Termination Formula]
\label{def:conditional-termination-formula}
 \begin{align*}
  \exists R, V \cdot \forall x, x' \cdot & I(x) \wedge L(x) \rightarrow V(x) ~ \wedge \\
                                 & L(x) \wedge V(x) \wedge T(x, x') \rightarrow V(x') \wedge R(x) > 0 \wedge R(x) > R(x')
 \end{align*}
\end{definition}

\begin{definition}[Multiple Loops Termination Formula]
\label{def:multi-termination-formula}
 \begin{align*}
  \exists R, V_1, V_2 \cdot \forall x, x' \cdot & I(x) \rightarrow V_1(x) ~ \wedge \\
                                        & V_1(x) \wedge L_1(x) \wedge T(x, x') \rightarrow V_1(x') ~ \wedge \\
                                        & V_1(x) \wedge \lnot L_1(x) \wedge L_2(x) \rightarrow V_2(x) ~ \wedge \\
                                        & V_2(x) \wedge L_2(x) \wedge T(x, x') \rightarrow V_2(x') \wedge R(x) > 0 \wedge R(x) > R(x')
 \end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula]
\label{def:nonterm-formula}
 \begin{align*}
  \exists N, x_0 \cdot \forall x \cdot \exists x' \cdot & I(x_0) \wedge L(x_0) \wedge N(x_0) ~ \wedge \\
							& N(x) \wedge L(x) \rightarrow T(x, x') \wedge N(x') \wedge L(x')
 \end{align*}
\end{definition}

\begin{definition}[Deterministic Non-Termination Formula]
\label{def:deterministic-nonterm-formula}
 \begin{align*}
  \exists N, x_0 \cdot \forall x, x' \cdot & I(x_0) \wedge L(x_0) \wedge N(x_0) ~ \wedge \\
							& N(x) \wedge L(x) \wedge T(x, x') \rightarrow N(x') \wedge L(x')
 \end{align*} 
\end{definition}
\end{framed}
\end{figure*}

@@

\begin{definition}[Synthesis Formula]
 \label{def:synth-formula}
 $$\exists P, x \cdot \forall y \cdot \sigma(P, x, y)$$
\end{definition}

@@

$$O(2^{K(f)}$$

@@

(Discussion of linearity vs. LKC here)

>>>>

\begin{definition}[Synthesis Fragment]
 A formula is in the synthesis fragment iff it is of the form
 \[
  \exists P, x . \forall y . \sigma(P, x, y)
 \]

@@

\begin{figure*}
\begin{framed}
\begin{definition}[Second-order Termination Formula]
\label{def:termination-formula}
\begin{align*}
 \exists R . \forall x, x' . & L(x) \wedge T(x, x') \rightarrow R(x) > 0 \wedge R(x) > R(x')
\end{align*}
\end{definition}

\begin{definition}[Conditional Termination Formula]
\label{def:conditional-termination-formula}
 \begin{align*}
  \exists R, V . \forall x, x' . & I(x) \wedge L(x) \rightarrow V(x) ~ \wedge \\
                                 & L(x) \wedge V(x) \wedge T(x, x') \rightarrow V(x') \wedge R(x) > 0 \wedge R(x) > R(x')
 \end{align*}
\end{definition}

\begin{definition}[Multiple Loops Termination Formula]
\label{def:multi-termination-formula}
 \begin{align*}
  \exists R, V_1, V_2 . \forall x, x' . & I(x) \rightarrow V_1(x) ~ \wedge \\
                                        & V_1(x) \wedge L_1(x) \wedge T(x, x') \rightarrow V_1(x') ~ \wedge \\
                                        & V_1(x) \wedge \lnot L_1(x) \wedge L_2(x) \rightarrow V_2(x) ~ \wedge \\
                                        & V_2(x) \wedge L_2(x) \wedge T(x, x') \rightarrow V_2(x') \wedge R(x) > 0 \wedge R(x) > R(x')
 \end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula]
\label{def:nonterm-formula}
 \begin{align*}
  \exists N, x_0 . \forall x . \exists x' . & I(x_0) \wedge L(x_0) \wedge N(x_0) ~ \wedge \\
							& N(x) \wedge L(x) \rightarrow T(x, x') \wedge N(x') \wedge L(x')
 \end{align*}
\end{definition}

\begin{definition}[Deterministic Non-Termination Formula]
\label{def:deterministic-nonterm-formula}
 \begin{align*}
  \exists N, x_0 . \forall x, x' . & I(x_0) \wedge L(x_0) \wedge N(x_0) ~ \wedge \\
							& N(x) \wedge L(x) \wedge T(x, x') \rightarrow N(x') \wedge L(x')
 \end{align*} 
\end{definition}
\end{framed}
\end{figure*}

@@

\begin{definition}[Synthesis Formula]
 \label{def:synth-formula}
 $$\exists P, x . \forall y . \sigma(P, x, y)$$
\end{definition}

@@

$$O\left(2^{K(f)}\right)$$

@@

\begin{proof}
 (Do this proof properly).

 Otherwise, we'd need to have an exponential number of programs computing the same function.
 That would be an unreasonable program encoding.
\end{proof}

Therefore the number of non-linear functions $f$ with $K(f) \leq \kappa$
grows exponentially with $\kappa$, while the number of linear functions
grows only linearly.  This shows that assuming the existence of a linear
ranking function is much stronger than assuming the existence of a short
termination proof.  In addition to being a much weaker assumption,
we argue that low Kolmogorov complexity is a more natural assumption than
linearity.  Humans tend to write programs they can understand.  It's hard
to quantify exactly what ``understandable'' means, but we think that
``having a short proof'' is closer to the mark than ``having a linear proof''.
